### 2018-04-16 concepts.tex
in progress

Many \emph{idempotent} operations fall into this category.
Consider adding an element to an AVL tree implementing a set, not a
multiset.  Assume the method call returns the parent node of the
element, whether it was added or was already in the tree.  Calling
this mehtod any number of times in sequence should always return the
same value, though the first call may have modified the contents of
the AVL tree.

@@

@@

Delta-debugging \cite{DD}\footnote{We use delta-debugging to stand for
  all test reduction algorithms, even those \cite{CReduce,tstl} that do not use the exact
  binary-search that distinguishes delta-debugging \emph{per se}; the
  differences are immaterial for our purposes.}  is a widely used method for reducing the
size of failing tests, making them easier to understand and debug.
Delta-debugging in the context of detecting nondeterminism has two
purposes.  One is simply the usual goal of reducing the size of a
test.  Identifying the cause of nondeterminism may be very easy in a
test consisting of ten library function calls (it is one of these ten
calls), but very difficult in a test consisting of a hundred library
function calls.  This is no different than the common use of
delta-debugging.  However, in nondeterminism detection,
delta-debugging can also be used to \emph{change the probability of
  nondeterministic behavior}.

In \emph{monotonic} cases, removing a part of a test cannot
increase the probability of a predicate holding.  This is fairly
common.  In these cases, when we use a reduction algorithm to reduce $t$
to $r$, $P(p(r) \leq P(p(t))$.  In \emph{non-monotonic} cases,
however, there is no such upper bound.  Removing a step in $t$ may
increase the probability that $t$ behaves nondeterministically.

>>>>

Many \emph{idempotent} operations fall into this category.
Consider adding an element to an AVL tree implementing a set, not a
multiset.  Assume the method call returns the parent node of the
element, whether it was added or was already in the tree.  Calling
this method any number of times in sequence should always return the
same value, though the first call may have modified the contents of
the AVL tree.

One application of vertical determinism, then, is to identify
idempotent operations in an automated test generation harness, and
simply automatically retry all such operations, checking that the
result is unchanged.  The overhead of vertical nondeterminism detection will
generally be much lower than that for horizontal nondeterminism
(horizontal checks must re-execute an entire test; vertical checks
only re-execute proportional to the number of idempotent operations
performed).  However, identifying idempotent operations is a
fairly serious specification burden.  Are there instances where a tool
can automatically identify a limited kind of idempotent behavior?

@@

Many library APIs are largely failure deterministic.  For instance, if
we exclude actual I/O errors, most POSIX file system calls either
succeed or do nothing (interfaces that may only partially succeed,
such as {\tt read} tend to explicitly return the degree of success
rather than signaling an error on partial success).  In fact, the
design of the POSIX {\tt write} call is carefully crafted to largely maintain
failure determinism.  If the failure mode is one that can be
anticipated, such as insufficient space to write all bytes requested
(but some bytes can be written), the call returns the actual bytes
written, and does not set a failure code.

In languages with a clear mechanism for expressing failure of a call
(e.g., exceptions in Python and Java, or Option/Result types in Rust,
Haskell, and ML), failure determinism can be automatically checked.
Moreover, the expected overhead should be even lower than the general
overhead for vertical determinism, in that we can expect most failing
operations to be fast in an automated test generation setting (that the
parameters to a call are illegal is the most common cause for
failure).

@@

Delta-debugging \cite{DD}\footnote{We use delta-debugging to stand for
  all test reduction algorithms, even those \cite{CReduce,tstl} that do not use the exact
  binary-search that distinguishes delta-debugging \emph{per se}; the
  differences are immaterial for our purposes.}  is a widely used method for reducing the
size of failing tests, making them easier to understand and debug.
Delta-debugging in the context of detecting nondeterminism has two
purposes.  One is simply the usual goal of reducing the size of a
test.  Identifying the cause of nondeterminism may be very easy in a
test consisting of ten library function calls (it is one of these ten
calls), but very difficult in a test consisting of a hundred library
function calls.  This is no different than the common use of
delta-debugging.  However, in horizontal nondeterminism detection,
delta-debugging also tends to \emph{change the probability of
  nondeterministic behavior}; this can be both harmful and
beneficial.  The behavior is part of a more general (and, to our
knowledge, not previously investigate) issue:  using delta-debugging
to minimize a test with respect to a predicate that only holds with a
certain probability (the predicate itself is nondeterministic).

In \emph{monotonic} probabilistic delta-debugging, removing a part of a test cannot
increase the probability of the predicate holding.  This is fairly
common.  In these cases, when we use a reduction algorithm to reduce $t$
to $r$, $P(p(r) \leq P(p(t))$.  In \emph{non-monotonic} cases,
however, there is no such upper bound.  Removing a step in $t$ may
increase the probability that $t$ behaves nondeterministically.

### 2018-04-16 experiments.tex
start on redis
in progress

\subsection {Case Study: Redis-Py}

@@

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  Running this harness using TSTL's {\tt
  --checkProcessDeterminism} flag required less than 10 seconds on
average to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm).  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue 
required another 92 seconds.

@@

>>>>

\subsection {Case Study: {\tt redis-py}}

The {\tt redis-py} \cite{redispy} module implements a widely-used Python interface
to the popular Redis in-memory database, cache, and
message broker \cite{redis}.

Using TSTL's harness for {\tt redis-py}, both {\tt redis-py} and Redis
itself can be tested; unfortunately, generating standalone
high-coverage regression tests for {\tt redis-py} and Redis using TSTL
was difficult, as numerous Redis commands introduce nondeterministic
behavior.  Some of these are obvious (e.g., {\tt randomkey, srandmember}) by
name.  Given an understanding of Redis semantics, that the various
commands producing data with a limited lifetime (e.g., {\tt expire,
  pexpire}) also introduce timing-based nondeterminism is also clear.
However, that a command such as {\tt restore} takes an expiration
argument is not obvious on inspection, nor is the behavior of {\tt
  spop} which pops a random value from a set.  Moreover, it is
difficult to guess whetehr the {\tt pipe} mechanism, which allows a
large sequence of commands to be queued up in a pipe and executed all
at once, introduced the potential for nondeterminism (does it execute
sequentially before other command are handled, or is it in parallel
with further commands).  Deep experience with Redis would make these
issues clear, but tests using a library are often written by those not
intimately familliar with the semantics of every used library (otherwise they would not have
been as likely to introduce flaky tests in the first place).  This is
even more true in the case of automated test generation, where the
test engineer is often chosen for expertise in test generation tools,
not in the domain of the software under test (SUT), and is seldom the
original developer of the code.

Using the original TSTL harness, approxoimately \emph{over 20\%} of all {\tt redis-py}
regression tests (of length 200) generated were flaky.  Using the
nondeterminism checker to reduce flaky tests to a minimal set allows
an engineer to leave an overnight run that will, in 12 hours, produce
a set of minimal flaky tests exposing the sources of
nondeterminism/flakiness.  In this case, the probability of flakiness
is not important, just the possibility, so we let the delta-debugging
create a minimal test with some non-zero (but possibly very small)
observed probability of nondeterminism.  The overhead for adding
nondeterminism checking is

After removing all identified sources of nondeterminism (the {\tt random-} functions, all
expiration-related functionality and {\tt restore} calls with non-zero
lifetimes, and {\tt spop}), but leaving all other suspect funcitonality in
place (e.g. pipes), no flakiness was observed in a sample of over 1,000 length
200 tests.  Moreover, because the removals were limited to acutally
observed sources of flakiness, the code coverage loss was minimal.
Average branch coverage, for regression tests of length 200 is reduced
by less than 5 branches.  Obviously, it is
impossible to test the removed calls, but the overall coverage loss is
both minimal and known, and can be made up for using specially-crafted
tests (for instance, wrapping the problematic calls in a way that does
not check the values, or adding a delay after an expiration to allow
the data to expire).  As a price to pay for the ability to produce
fast-executing high-coverage full regression tests (not just tests for
crashes and unexpected exceptions), this is a very acceptable price.
Moreover, turning off sources of known, expected nondeterminism makes
it possible to aggressively test {\tt redis-py} and Redis for
unexpected nondeterminism arising from actual bugs.  The overhead for such
determinism checks, with a 0 second delay between operations, is only
20\%, much lower than the expected cost of running each test twice.
This is because choosing the actions in a test (and determining which
actions are enabled at each step) consumes a large part of the test
generator's time; running the test again and checking equality is
relatively inexpensive.

Figure \ref{hashbug} shows TSTL code for generating inputs to the {\tt
  datarray} algorithms.  Running this harness using TSTL's {\tt
  --checkProcessDeterminism} flag required less than 10 seconds on
average to produce a test exhibiting process-level nondeterminism in
the {\tt calc\_marginals\_sumproduct} function (the only broken
algorithm).  Reducing this 60 step test to a minimal test of only 6 steps,
showing an extremely simple input producing the issue, 
required another 92 seconds.

Removing the nondeterministic call, we can see that the cost of
checking for process nondeterminism, with a zero delay on operations,
is high, a 93\% slowdown.  This is due to the very high cost of
using a new subprocess to re-run the test, and comparing values using
text I/O on output pipes, rather than on values alone.

@@

@@

\subsection {Vertical Determinism Case Study: {\tt pyfakefs}}
